---
title: "Reproducible Research: Project 1"

---
### By Amit Morathkar


## Loading and preprocessing the data

```{r loaddata}
unzip(zipfile="activity.zip")
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
library(ggplot2)
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```

## What is the average daily activity pattern?
```{r}
library(ggplot2)
avg <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=avg, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken")
```

```{r}
avg[which.max(avg$steps),]
```
## Imputing missing values
Values coded with NA
```{r Nos_missing}
missing <- is.na(data$steps)
# How many missing
table(missing)
```
# Replace missing values with mean value of its 5 min interval
```{r}
add.value <- function(steps, interval) {
    added <- NA
    if (!is.na(steps))
        added <- c(steps)
    else
        added <- (avg[avg$interval==interval, "steps"])
    return(added)
}
added.data <- data
added.data$steps <- mapply(add.value, added.data$steps, added.data$interval)
```
```{r}
total.steps <- tapply(added.data$steps, added.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps)
median(total.steps)
```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
added.data$date <- as.Date(added.data$date)
added.data$day <- sapply(added.data$date, FUN=weekday.or.weekend)
```

Panel plot containing plots of average number of steps taken
on weekdays and weekends.
```{r}
avg <- aggregate(steps ~ interval + day, data=added.data, mean)
ggplot(avg, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
    xlab("5-minute interval") + ylab("Number of steps")
```